--- 
# title: Literature review
coverpage: true
toc: true
author: "François Leroy, PhD student at CZU"
date: "`r Sys.Date()`"
output: pdf_document
documentclass: report
classoption: oneside
fontsize: 12pt
linestretch: 1.5
geometry:
- left = 2cm
- right = 2cm
- top = 2cm
- bottom = 2cm
colorlinks: yes
link-citations: yes
bibliography:
 - references.bib
biblio-style: apa
# biblio-title:  ## note: the name and label of the bib part is in the template!!!
---

```{r, echo = F}
knitr::opts_chunk$set(warning = F,
                      message = F)
```

```{r, echo=F}
library(bibliometrix)
```


# Outline {-}

Literature review about the link between biodiversity facets trends and spatial/temporal scales. 

The idea is to take every paper that talk about biodiversity trends (so far using just the species richness seems already a lot of paper) and to list **1)** which biodiversity metric they use **2)** which taxon/taxa they use, **3)** the spatial scale, **4)** the temporal scale and **5)** what is the dynamic (does the biodiversity metric increase/decrease/doesn't change over time/unclear).

Make a table of all these papers and `group_by(taxa) %>% order_by(spatial_scale | temporal_scale)`. Then see if for each taxa we can find a trend (a bit like in Chase *et al.* 2019 Oikos paper | Jarzyna *et al.* 2015 but here I am not making the analysis, just taking the analysis from papers). Best example found so far: Hill & Hamer 2004 (https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.0021-8901.2004.00926.x)

I am using the "Advanced Research" tab of Web of Science which allows me skim through the entire literature using a convenient syntax. For instance:

```
AB = ((biodiversity OR species richness OR diversity) AND
(temporal trend* OR dynamic*) AND
(bird* OR avia*)) 
```

From this code, I could change the taxon.


# Introduction

@dornelas_assemblage_2014 

# References {-}
\singlespacing
