--- 
# title: Literature review
coverpage: true
toc: true
author: "François Leroy, PhD student at CZU"
date: "`r Sys.Date()`"
output: pdf_document
documentclass: report
classoption: oneside
fontsize: 12pt
linestretch: 1.5
geometry:
- left = 2cm
- right = 2cm
- top = 2cm
- bottom = 2cm
colorlinks: yes
link-citations: yes
bibliography:
 - references.bib
biblio-style: apa
# biblio-title:  ## note: the name and label of the bib part is in the template!!!
---

```{r, echo = F}
knitr::opts_chunk$set(warning = F,
                      message = F)
```

```{r, echo=F}
library(bibliometrix)
```


# Outline {-}

Literature review about the link between biodiversity facets trends and spatial/temporal scales. 

The idea is to take every paper that talk about biodiversity trends (so far using just the species richness seems already a lot of paper) and to list **1)** which biodiversity metric they use **2)** which taxon/taxa they use, **3)** the spatial scale, **4)** the temporal scale and **5)** what is the dynamic (does the biodiversity metric increase/decrease/doesn't change over time/unclear).

Make a table of all these papers and `group_by(taxa) %>% order_by(spatial_scale | temporal_scale)`. Then see if for each taxa we can find a trend (a bit like in Chase *et al.* 2019 Oikos paper | Jarzyna *et al.* 2015 but here I am not making the analysis, just taking the analysis from papers). Best example found so far: [Hill & Hamer 2004](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.0021-8901.2004.00926.x)

I am using the "Advanced Research" tab of Web of Science which allows me skim through the entire literature using a convenient syntax. For instance:

```
AB = ((biodiversity OR species richness OR diversity) AND
(temporal trend* OR dynamic*) AND
(bird* OR avia*)) 
```
And

```
AB = ((biodiversity change index)  AND (bird*  OR avia*)  AND trend*)
```

And

```
AB = ((species richness) AND (bird* OR avia*) AND trend*) 
```


From this code, I could change the taxon.


| Article  	| Metric    | Spatial scale	| Temporal scale | Trend 	| Location |  
|---	|---	|---  |---	|---	|---  |
|   	|   	|   	|   	|   	|     |

# Dashboard {-}

[Reference paper](https://www.sciencedirect.com/science/article/pii/S1470160X20306658?via%3Dihub)

* 05/07/2021: research wos made with the literature review filter for the first query (stopped at #13) and created the second query (stopped at #2)

* 07/07/2021: questions to Petr: **1)** can the geometric mean of relative abundance + the weighted goodness of fit be used as biodiversity trend index, **2)** can the Farmland Bird Indicator (FBI) be used as biodiversity trend (for me it is more biodiversity health, Chiron et al 2013) **3)** what about the Red List Index trend? **4)** what about Multispecies population indexes?

**The question could be: do I look also to the trend of qualitative index of biodiversity?**

* 08/07/2021: stopped at the article 41 for research #2. 


# Introduction

Human life quality is intrinsically linked to ecosystems state that he is living in. Indeed, ecosystems services extend in a large spectrum of mechanisms including nutrient cycle, food production, or climate and water cycle regulation [@pereira_global_2012]. Some of those ecosystem functions are managed by bird biodiversity such as seed dispersal, controls pests or pollinate plant. Unfortunately, anthropogenic stressors like habitat loss, over exploitation, pollution or introduction of invasive species could lead biodiversity to its sixth mass extinction [@barnosky_has_2011].

Biodiversity erosion is now known from everyone and political decisions has been stated in order to limit it [*e.g.* @the_convention_on_biological_diversity_convention_2021, 2010, 2002]. However, these objectives have been so far not reached due mainly to our confusion and misunderstanding about biodiversity dynamic and how to determine it.

Studying biodiversity can be confusing, especially because several choices must be done. Firstly, the level at which you are looking at the biodiversity must be chosen (*e.g.* species, functional, phylogenetic diversity). Secondly, one must decide which metric is the most appropriate for his study. There are many facets of biodiversity that can be measured by different metrics depending on the objective of your study. Measures of static biodiversity are commonly used such as species richness or $\alpha$ diversity [*i.e.* number of species, @whittaker_vegetation_1960], the Shannon index [@shannon_mathematical_1948] ,the Simpson index [@simpson_measurement_1949] or the Hill number [@hill_diversity_1973]. The later three biodiversity indexes take into account the relative abundances of the species and can be considered as the *quality* of the biodiversity. On an other hand, the spatial and temporal $\beta$ diversity will measure the species turnover and can be measured thanks to Whittaker's [@whittaker_evolution_1972], Sørensen's [@sorensen_method_1948] or Jaccard's [@jaccard_distribution_1912] dissimilarity indexes [*e.g.* @keil_patterns_2012].

However, overall biodiversity (*i.e.* taking into account species of every taxa) may not be relevant for one's case study. Thus, several multi-species indicators have also been created, taking into account the abundances of indicator species giving information on the ecosystem health. The most known ones are the Red List Index [@butchart_improvements_2007; @butchart_using_2005; @butchart_measuring_2004] or the Biodiversity Change Index [@normander_indicator_2012]. (talk about the geometric mean and GoF)


**Add a part about multispecies indicators** Freixedas et al.

**Add a part about the trends of these metrics inspired from just below**

While the loss of global biodiversity is unprecedented, current scientific literature has also shown that temporal trends in local changes of biodiversity can be opposite to trends at larger scales [@chase_species_2019]. Thus, current changes in biodiversity is far more complex than a simple global decrease: most of the ecosystems undergo alterations of their communities with changes in species composition [@blowes_geography_2019; @dornelas_quantifying_2013].

# References {-}
\singlespacing
