---
# title: Literature review
coverpage: true
toc: true
author: "François Leroy, PhD student at CZU"
date: "`r Sys.Date()`"
documentclass: report
classoption: oneside
fontsize: 12pt
linestretch: 1.5
geometry:
- left = 2cm
- right = 2cm
- top = 2cm
- bottom = 2cm
colorlinks: yes
link-citations: yes
bibliography: references.bib
biblio-style: apa # authoryear
---

```{r, echo = F}
knitr::opts_chunk$set(warning = F,
                      message = F,
                      echo = F)
```

```{r, echo=F}
library(bibliometrix)
library(googlesheets4)
library(dplyr)
library(kableExtra)
library(ggplot2)
library(kableExtra)
library(tidyr)
library(forcats)
```

<!-- # Outline {.unnumbered} -->

<!-- Literature review about the link between biodiversity facets trends and spatial/temporal scales. -->

<!-- <!-- The idea is to take every paper that talk about biodiversity trends (so far using just the species richness seems already a lot of paper) and to list **1)** which biodiversity metric they use **2)** which taxon/taxa they use, **3)** the spatial scale, **4)** the temporal scale and **5)** what is the dynamic (does the biodiversity metric increase/decrease/doesn't change over time/unclear). -->

<!-- Make a table of all these papers and `group_by(taxa) %>% order_by(spatial_scale | temporal_scale)`. Then see if for each taxa we can find a trend (a bit like in Chase *et al.* 2019 Oikos paper \| Jarzyna *et al.* 2015 but here I am not making the analysis, just taking the analysis from papers). Best example found so far: [Hill & Hamer 2004](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.0021-8901.2004.00926.x) -->

<!-- **Meeting with Petr** -->

<!-- 1)  How is the trend of biodiversity metrics linked to spatial grain/spatial extent and temporal grain/temporal extent. -->

<!-- -   Talk about papers which does compare (Jarzyna Jetz) -->
<!-- -   The inconsistency of reporting scales (especially for models as in Jiguet et al.or Chiron et al., and for MSI metrics) -->

<!-- 2)  Data heterogeneity -->

<!-- -   Lack of spatial replication. That's why this review is important -->
<!-- -   The inconsistency of reporting scales (especially for models, and for MSI metrics) -->
<!-- -   Lack in other than western countries -->
<!-- -   Metric heterogeneity (Freixedas 2001 but they don't mention the macroecological ones as McGill, my struggle because of the time for space substitution as Hill & Hammer). Some metrics takes into account temporal dynamic but some papers don't look at the trend of these metrics (Jarzyna et al. 2015) -->
<!-- -   Can we use space for time substitution as an actual substitution (Hill & Hammer) -->
<!-- -   Pop trends are usually stronger (cf. all the abundance metrics) -->

<!-- 3)  Future directions -->

<!-- -   Lack of data for other thsn western cvountries: important to harvest data in other countries -->

<!-- # Dashboard {.unnumbered} -->

<!-- [Reference paper](https://www.sciencedirect.com/science/article/pii/S1470160X20306658?via%3Dihub) -->

<!-- -   05/07/2021: research wos made with the literature review filter for the first query (stopped at \#13) and created the second query (stopped at \#2) -->

<!-- -   07/07/2021: questions to Petr: **1)** can the geometric mean of relative abundance + the weighted goodness of fit be used as biodiversity trend index, **2)** can the Farmland Bird Indicator (FBI) be used as biodiversity trend (for me it is more biodiversity health, Chiron et al 2013) **3)** what about the Red List Index trend? **4)** what about Multispecies population indexes? -->

<!-- -   08/07/2021: stopped at the article 41 for research \#2. -->

<!-- -   12/08/2021: stopped at article 4 for research \#4 -->

<!-- -   13/08/2021: stopped at article 8 for research \#4 -->

<!-- -   17/08/2021: stopped at article 15 for research \#4 -->

<!-- -   18/08/2021: stopped at article 30 for research \#4 -->

<!-- -   19/08/2021: stopped at article 46 for research \#4 -->

<!-- -   20/08/2021: stopped at article 64 for research \#4 -->

<!-- -   01/09/2021: verifying spatial scales stopped at Dittrich 2019 -->

<!-- -   02/09/2021: **Question 1:** for the FBI/WBI...\*BI indexes, usually they use a GLM/GAM to predict the abundance over the entire spatial extent and then compute the metric. Basically, those metrics are geometric means of predicted species abundances. Which spatial scale to use: the spatial unit of the prediction (i.e. the plot), or the entire area predicted? (Imo, the second option is correct). **Question 2:** same question for the Geometric mean but I am realizing while writing this question that \*BI are kind of similar to geometric means so answering the first question will answer this one. -->

<!-- **Papers that are driving me mad:** @doxa_low-intensity_2010, @jiguet_french_2012, and @chiron_forecasting_2013, @eglington_disentangling_2012 -->

# Introduction

Human life quality is intrinsically linked to ecosystems state that he is living in. Indeed, ecosystems services extend in a large spectrum of mechanisms including nutrient cycle, food production, or climate and water cycle regulation [@pereira_global_2012]. Some of those ecosystem functions are managed by bird biodiversity such as seed dispersal, controls pests or pollinate plant. Unfortunately, anthropogenic stressors like habitat loss, over exploitation, pollution or introduction of invasive species could lead biodiversity to its sixth mass extinction [@barnosky_has_2011].

<!-- Biodiversity erosion is now known from everyone and political decisions has been stated in order to limit it [*e.g.* @the_convention_on_biological_diversity_convention_2021, 2010, 2002]. However, these objectives have been so far not reached due mainly to our confusion and misunderstanding about biodiversity dynamic and how to determine it. -->

We now know that the loss of global biodiversity is unprecedented and political decisions has been stated in order to limit it [*e.g.* @secretariat_of_the_convention_on_biological_diversity_global_2006, 2010, 2002]. However, current scientific literature has also shown that temporal trends in local changes of biodiversity can be opposite to trends at larger scales [*e.g.* @chase_species_2019]. Thus, current changes in biodiversity is far more complex than a simple global decrease: most of the ecosystems undergo alterations of their communities with changes in species composition [@blowes_geography_2019; @dornelas_quantifying_2013]. Wonders persist about how the trend of these different metrics of biodiversity are link to the spatial and temporal scales used when measured.

In order to investigate this link between spatio-temporal scales and biodiversity metrics, birds is a relevant taxon. Thanks to the many ornithological monitoring and surveys, we now have a large number of long, high-quality time series on bird populations [@bejcek_velke_2016; @sauer_north_2013; @kamp_population_2021...]. Birds are easy to observe, easy to identify and thus many volunteers are motivated to conduct standardized sampling. Given their ability to change quickly of locations, their presence is also a good indicator for ecosystem health and thus several standardized metrics have been created to assess their populations.

However, studying biodiversity can be confusing, especially because several choices must be done. Firstly, the level at which you are looking at the biodiversity must be chosen (*e.g.* taxonomic, functional, phylogenetic diversity). Secondly, one must decide which metric is the most appropriate for his study. There are many facets of biodiversity that can be measured by different metrics depending on the objective of your study. Measures of static biodiversity are commonly used such as species richness or $\alpha$ diversity [*i.e.* number of species, @whittaker_vegetation_1960], the Shannon index [@shannon_mathematical_1948] ,the Simpson index [@simpson_measurement_1949] or the Hill number [@hill_diversity_1973]. The later three biodiversity indexes take into account the relative abundances of the species and can be considered as the *quality* of the biodiversity. On an other hand, the spatial and temporal $\beta$ diversity will measure the species turnover and can be measured thanks to Whittaker's [@whittaker_evolution_1972], Sørensen's [@sorensen_method_1948] or Jaccard's [@jaccard_distribution_1912] dissimilarity indexes [*e.g.* @keil_patterns_2012]. All these metrics assess the taxonomic diversity, *i.e.* they use the species as unit. However, it has also been shown that functional and phylogenetic diversity can provide supplementary information on the community structure and its dynamic [*e.g.* @mcgill_rebuilding_2006; @mouquet_ecophylogenetics_2012; @webb_phylogenies_2002].

An other measure of biodiversity of great interest is the abundance of the populations. As a matter of fact, as individuals react to disturbances of ecosystems by disappearing, the population trends are considered as resulting of the ecosystems health. Overall populations are impossible to assess, but determining the abundance of few indicator species can give a relevant summary on how is going the entire ecosystem [@gregory_developing_2005]. These family of metrics are called the multi-species indicators (MSI) and we have seen the emergence of several of them according to the ecosystem considered. We can cite here the farmland bird indicator, woodland bird indicator or one of the most informative, which summarizes these two previous metrics: the Wildland Bird Indicator [@gregory_generation_1999; @gregory_wild_2010]. The main idea of these metrics is to compute the geometric mean of abundance of few key species over time.

Finally a last class of indicators, not considered here, take into account both species and ecosystems feature into a summarizing index. The most known ones are the Red List Index [@butchart_improvements_2007; @butchart_using_2005; @butchart_measuring_2004] or the Biodiversity Change Index [@normander_indicator_2012].

<!-- For instance, the geometric mean of relative abundances or the goodness-of-fit statistic [@studeny_goodness_2011] are some of the baseline. Other multi-species indicators have also been created specifically for birds, such as the Farmland Bird Indicator [@gregory_developing_2005], the Forest Bird Indicator [@gregory_population_2007] or the Wild Bird Indicator [@gregory_wild_2010]. -->

Here, I propose to review articles assessing the temporal trends of different macroecological avian biodiversity metrics and to look at which spatial and temporal scales these studies have been done. I decided to consider the most common indicators used to assess biodiversity at the community level and higher, such as the diversity indexes (*e.g.* species richness, functional diversity...) or the abundance indexes [@mcgill_fifteen_2015]. Summarizing the trends of these qualitative and/or quantitative avian biodiversity indexes along with their spatial and temporal scales will help to see more clearly how the trends of biodiversity are linked to spatio-temporal scales. It is also important to demonstrate that the information about the sampling plan (*i.e.* spatial scale, time span, temporal scales etc) is not systematically indicated in the scientific literature and can bring confusion to the analysis and comparisons of their trends. I believe that this review can help to have a better overview of the current knowledge on the link between trends of biodiversity metrics of bird populations and the spatio-temporal scales.

\textcolor{red}{specify that it is mainly for continental/coastal birds but no look at  islandic communities and change *Stable* to *No trend*}

# Materials and Methods

```{r, echo=FALSE}
## Load the table
tab <- read_sheet("https://docs.google.com/spreadsheets/d/1SSvav_ISY1h0N7zr6xgxSyrQV8_XpdpsIUYEWzikaCA/edit#gid=0")

## Discretizing spatial scales
tab <-
  tab %>%
    mutate(`Spatial grain (Km²)` = as.character(`Spatial grain (Km²)`)) %>% 
    mutate(`Spatial grain (Km²)` = case_when(
      `Spatial grain (Km²)` <= 25 | `Spatial grain (Km²)` == "Local" ~ "Local",
      (`Spatial grain (Km²)` > 25 & `Spatial grain (Km²)` < 800000) | (`Spatial grain (Km²)`== "Admin. unit") ~ "Regional",
      (`Spatial grain (Km²)` == "National") | (`Spatial grain (Km²)`== "National unit") | (`Spatial grain (Km²)`== "Europe") ~ "National",
      TRUE ~ "Global"
    )) 
```

For this review, articles of interest were the ones assessing temporal trends of the most common indicators (*i.e.* metrics) of avian biodiversity and specifying spatial and temporal scales. For this, I used the *"advanced search"* tool of the ISI Web of Science Core collection database with these four following queries:

1.  `AB = ((biodiversity OR species richness OR diversity) AND (temporal trend* OR dynamic*) AND (bird* OR avia*))` which resulted in 1346 references.

2.  `AB = ((biodiversity change index)  AND (bird*  OR avia*)  AND trend*)` which resulted in 60 references.

3.  `AB = ((species richness) AND (bird* OR avia*) AND trend*)` which resulted in 313 references.

4.  `ALL=(birds AND species richness AND temporal trend)` which resulted in 88 references.

For each query, the title and abstract of the articles were reviewed. When the temporal trend was explicitly specified (either visually or literally), the material and method part was read in order to collect the *spatial grain* of the trend (*i.e.* the area at which the trend is assessed), its *temporal grain* (*i.e.* the time span at which data have been gathered at one census session), the *spatial extent* (*i.e.* the entire area at which the study applies), the *temporal extent* and the *beginning and ending years* of the study as well as the *general trend* of the metric (Table \@ref(tab:maintable)). Alternatively, the articles which were often referred to in the relevant articles were also explored.

I decided to take into account only the articles for which there was spatial replicates, *i.e.* where the trend of the metric was assessed at several locations with the same spatial grain. With these replications, the trend reported at one spatial grain is more reliable. However, at larger spatial grains (*i.e.* national, continental or global scales), spatial replicates are rarer. Thus, I decided to take note of these unreplicated trends only when they were found in a article were spatial replicates were used at smaller scales. 

Spatial grain sizes are discretized into four levels: *local* $<= 25$ $Km^2$, *regional* $> 25$ $Km²$, *national* when an entire country is considered and *global* at the worldwide scale.

Concerning the trend assessment, different papers contain the *p-value* or directly specify the significant trend of the metric. However, a portion of papers gives only visual representations of the trend. For those, the standard error was used when displayed. For the very few only giving the trend, \textcolor{red}{the rule of thumb was applied}. Information can be found in the column *Note* of the Table \@ref(tab:notetable) of the supplementary material. Moreover, the final trend retained (*i.e.* either *Increase*, *Stable* or *Decrease*) doesn't reflect all the fluctuations of the metric through time but rather the difference between the starting and ending points.

I have found `r nrow(unique(read_sheet("https://docs.google.com/spreadsheets/d/1SSvav_ISY1h0N7zr6xgxSyrQV8_XpdpsIUYEWzikaCA/edit#gid=0", sheet = 2)[,1]))` references in which authors were both determining the temporal trend of a metric of interest and explicitly defining the grain size. However, only `r length(unique(tab$Reference))` of them are using spatial replicates and are thus relevant for this study (Table \@ref(tab:maintable)). The classes of metric are: *Species richness, `r paste(as.vector(unique(tab$Metric)[-2]), collapse = ", ")`*. Some of this classes contain several different indexes. This is the case for instance for the class *Diversity*, which can be defined by either the Shannon or the Simpson index, or for the class *Abundance* which contains various multi-species indicators (for more information, see Table \@ref(tab:notetable)).

<!-- Moreover, @pilotto_meta-analysis_2020 conducted a meta-analysis in which they computed and summarized the trend of four biodiversity metrics (namely, species richness, species diversity, abundance and temporal turnover). Some of them were concerning bird communities. For those latter, I used their code and data on the [github repository](https://github.com/FrsLry/R-code) of their paper in order to compute the trends of these four metrics for the bird datasets. -->

```{r maintable, echo=F}
# Function to collapse the reference column
collapse_rows_df <- function(df, variable){

  group_var <- enquo(variable)

  df %>%
    group_by(!! group_var) %>%
    mutate(groupRow = 1:n()) %>%
    ungroup() %>%
    mutate(!!quo_name(group_var) := ifelse(groupRow == 1, as.character(!! group_var), "")) %>%
    select(-c(groupRow))
}


tab %>%
  # mutate(`Spatial grain (Km²)` = as.numeric(`Spatial grain (Km²)`),
  mutate(
         `Temporal grain (year)` = as.numeric(`Temporal grain (year)`),
         `Spatial extent (Km²)` = as.numeric(`Spatial extent (Km²)`),
         `Temporal extent (year)` = as.numeric(`Temporal extent (year)`)) %>%
  select(-Note) %>%
  as_tibble() %>%
  group_by(Reference) %>%
  collapse_rows_df(Reference) %>%
  ## escape = F is very important for the citation
  kbl(booktabs = T, escape = F, format = "latex", longtable = T,
      caption = "Trends of different metrics of biodiversity at various spatial and temporal scales") %>%
  kableExtra::kable_styling(
   latex_options = c("hold_position", "repeat_header", "striped"),
   font_size = 10,
   full_width = FALSE
   ) %>%
  column_spec(1:9, width = "6.5em") %>%
  landscape()
```


# On the link between spatial grain and trends of biodiversity

The median spatial extent of the `r length(unique(tab$Reference))` articles is `r round(median(as.numeric(unique(unlist(tab[,5])))))` $Km^2$, with the smallest area of `r min(as.numeric(unlist(tab[,5])))` $Km^2$ and the greatest representing the global emerged surface (*i.e.* `r round(max(as.numeric(unlist(tab[,5]))))` $Km^2$). Overall, there were `r table(tab$Trend)[1]` *Decrease*, `r table(tab$Trend)[2]` *Increase* and `r table(tab$Trend)[3]` *Stable* reliable trends (*i.e.* spatially replicated) across the literature, without consideration of any metric or spatial scales. 

One have to keep in mind that the proportions of the trends presented here is not representative of the overall the scientific literature. Taking only the articles with spatial replicates reduces the number of studies and thus doesn't reflect the overall dynamic of biodiversity, which is not the purpose of this review. As a matter of fact, meta-analysis by @fraixedas_state_2020 indicates that around half of bird related indicators are declining and that only $20\%$ of them are increasing.

In our case, local scales are more represented than the others and the number of papers decreases with the increasing spatial scale (Figure \ref{fig:barspatscale}). This is expected, as the spatial replications get more demanding in organization and resources as the grain size enlarges. The *Increase* of the metrics seems to be dominating at smaller scales. On an other hand, the proportion of *Decrease* is more important at regional scales than at local scales. At the global scale, no *Increase* were found.

```{r barspatscale, fig.cap="Proportion of *Increase*, *Decrease* or *Stable* trends for each spatial scale", fig.pos = "!H"}

table(tab$`Spatial grain (Km²)`, tab$Trend) %>% 
  as.data.frame() %>% 
  rename("Frequency" = Freq,
         "Spatial grain size" = Var1,
         "Trend" = Var2) %>% 
  mutate("Spatial grain size" = fct_relevel(`Spatial grain size`,c("Local", "Regional", "National", "Global"))) %>% 
  ggplot(aes(fill = Trend, x = `Spatial grain size`, y = Frequency))+
  geom_bar(position = "stack", stat = "identity")+
  theme_light()+
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.x=element_blank())+
  scale_fill_grey()

# tmp <- cbind(table(tab %>% select(Metric, Trend))/rowSums(table(tab %>% select(Metric, Trend))), n = rowSums(table(tab %>% select(Metric, Trend)))) %>% as_tibble(rownames = "rownames")
# 
# 
# kable(tmp %>%
#         mutate_at(c("Decrease", "Increase", "Stable"), funs(format(round(., 2), nsmall = 2)))) %>%
#   column_spec(1, width = "20em")
```


Replicated species richness and abundance metrics are very represented in the scientific literature (Figure \ref{fig:barmetrics}). The less common trend of abundance is *Increase*, whilst *Decrease* and *Stability* are both as common. Diversity indexes (*i.e.* Sørensen and Jaccard) were always found increasing and temporal $\beta$-diversity was found most of the time increasing and never decreasing. The small frequency of the other metrics doesn't allow one to discuss the trend proportions.

```{r barmetrics, fig.cap="Proportion of *Increase*, *Decrease* or *Stable* trends for each of the metric", fig.pos = "!H", }

table(tab %>% select(Metric, Trend)) %>% 
  as.data.frame() %>% 
  rename("Frequency" = Freq) %>% 
  ggplot(aes(fill = Trend, y = `Frequency`, x = Metric))+
  geom_bar(position = "stack", stat = "identity")+
  theme_light()+
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.x=element_blank())+
  scale_fill_grey()

# tmp <- cbind(table(tab %>% select(Metric, Trend))/rowSums(table(tab %>% select(Metric, Trend))), n = rowSums(table(tab %>% select(Metric, Trend)))) %>% as_tibble(rownames = "rownames")
# 
# 
# kable(tmp %>%
#         mutate_at(c("Decrease", "Increase", "Stable"), funs(format(round(., 2), nsmall = 2)))) %>%
#   column_spec(1, width = "20em")
```


**At local scales**, species richness experiences dominantly increase (Figure \ref{fig:barmetricsperspatscale}). Evenness indexes, *i.e.* taxonomic and functional evenness, are mainly found increasing. Concerning the abundance indexes, the stability dominates but the increase is also often witnessed. In contrast, **at regional scale**, abundance metrics always decrease, whilst temporal $\beta$-diversity always increase. Species richness experiments mainly increases, then stability but no decrease is reported. Concerning **national and global scales**, one have to keep in mind that spatial replicates are less frequent and that most of the trends reported here for these two spatial scales are not replicated. Exceptions are made by @bowler_geographic_2021 and @donald_agricultural_2001. The former showed negative trends in abundance indexes for Denmark and Germany and positive trends for Switzerland and Czech Republic, indicating that one can not conclude about general trend at national scale (here referred as *Stable*). For @donald_agricultural_2001, trends of mean populations were computed for 30 european countries and most of them were negatives. No positive trend seem dominant at these spatial scales. Global trends only come from @jarzyna_taxonomic_2018.

```{r barmetricsperspatscale, fig.cap="Proportion of *Increase*, *Decrease* or *Stable* trends for each metric. Each panel represent one spatial scale", fig.pos = "!H"}
table(tab %>% select(Metric, Trend, `Spatial grain (Km²)`)) %>% 
  as.data.frame() %>% 
  rename("Frequency" = Freq) %>% 
  filter(Frequency != 0) %>% 
  rename(`Spatial grain (Km²)` = `Spatial.grain..Km².`) %>% 
  mutate(`Spatial grain (Km²)` = fct_relevel(`Spatial grain (Km²)`,c("Local", "Regional", "National", "Global"))) %>% 
  ggplot(aes(fill = Trend, y = `Frequency`, x = Metric))+
  geom_bar(position = "stack", stat = "identity")+
  facet_wrap(~ `Spatial grain (Km²)`, nrow = 1, scales = "free_x")+
  theme_light()+
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 7),
        axis.title.x=element_blank())+
  scale_fill_grey()

``` 

Number of studies using spatial replicates is limited. Thus, observing general rules about spatial scaling of trends of macroecological metrics is not possible with the current state of the scientific literature. However, concerning species richness, we can see that the decrease is global, but that this decrease is rare at lower spatial scales. In fact, increases are  more often observed, confirming the high perturbations that biodiversity is undergoing [@dornelas_assemblage_2014; @vaidyanathan_worlds_2021]. This analysis goes along with the temporal $\beta$-diversity which is always observed either stable or increasing,  sign of instability of the ecosystems. 

On an other hand, the decrease of abundances seem to appear mainly at regional or national scales. \textcolor{red}{This may be due to the fact that local scale samplings are done usually outside of very anthropized  ecosystems, such as cities.Increasing the scale to regional and national increases the proportion of cities and farmland area, which can have negative impact on bird populations}  [@reif_long-term_2013].

# A lack of consistency

## Spatial replicates are not a standard

Articles computing the trend of a metric with spatial replicates were limited, either due to a lack of data or because the trend was assessed for the spatial extent of the data. For instance, the Breeding Bird Surveys [*e.g.* @sauer_north_2013; @kamp_population_2021] follow standardized sampling plan with spatial replications (*i.e.* census plots). However, not all the trend reported for the BBS are summarized at their specific grain sizes and were sometimes computed at their respective national scales, therefore without spatial replication. For instance, the common method encountered to assess population trends (*i.e.* abundance indexes) is to learn a predictive model from the data, predict the target feature (*i.e.* abundance) and then compute the metric and its trend from the output of the model at the national spatial extent [*e.g.* @jiguet_modeling_2005; @jiguet_french_2012; @eglington_disentangling_2012; @doxa_low-intensity_2010; @sauer_first_2017]. These analysis are essential from a conservation point of view and are a majority [@fraixedas_state_2020]: they give valuable information about ecosystem health at large spatial extent and are thus used as arguments by decision-makers. However, they represent just one trend amongst all the possible trends that can be found at this spatial scales and thus doesn't give information about the link between spatial grain and temporal trend. An other common type of studies are the one using the space-for-time substitution [@walker_use_2010] to assess the trend of a metric [one of the best example is @hill_determining_2004]. One could think that using this studies could increase significantly the spatial replicates. However, the space-for-time substitution is mainly used to assess the impact of a process (*e.g.* before/after logging, before/after urbanization etc) meaning that the trend computed is higly biased, which we try to avoid for our topic.      

Even less articles were computing the trends of metrics with spatial replicates at various spatial grain. This was the case for only @chase_species_2019 and @jarzyna_taxonomic_2018. It is also important to cite here @jarzyna_spatial_2015, who did iteration computing of temporal change community metrics (*i.e.* temporal dissimilarity, temporal turnover, extinction and colonization) at several spatial scales. However, the temporal trend of these metrics weren't considered and are therefore not reported in Table \@ref(tab:maintable).

## Metric heterogeneity

Macroecology has seen its metrics blooming these last decades, in search of a multiparametric indexes taking into account all the components of an ecosystem, as do the composite indicators [*e.g.* the red list index, @butchart_improvements_2007; biodiversity change indexes, @normander_indicator_2012]. The latter have proven very effective in order to assess conservation policies and are able to synthesize the ecosystem complexities. This has been especially the case for the birds, who are of great interest in assessing the dynamic of ecosystem health [@fraixedas_state_2020]. Thus, studies focusing only on macroecological metrics are less common than expected. Coupled with the need of spatial replicates, studying the link between the trend of macroecological metrics and spatio-temporal scales still needs to be explored.  

## On the temporal scale

The oldest study started in `r min(as.numeric(unlist(unique(tab$Years) %>% strsplit("-"))), na.rm = T)` and the median duration is `r round(median(as.numeric(unique(unlist(tab[,6]))), na.rm = TRUE))` years, with a minimum time span of `r min(as.numeric(unlist(tab[,6])), na.rm = TRUE)` years and a maximum of `r max(as.numeric(unlist(tab[,6])), na.rm = TRUE)`.

Determination of the temporal grain was way more complicated, as there is no consensus on the subject. Usually, the temporal grain of the sampling is specified, but sometimes with inaccuracies (*e.g.* "in the early morning"). However, the temporal grain of the sampling plan doesn't represent the final temporal grain. For instance, some metrics are summed over a certain spatial extent [e.g. summing the species richness over an atlas square, like in @van_turnhout_scale-dependent_2007]. In this case, the temporal grain should also be summed, but this was not specified. On an other hand, when the metric is averaged over a spatial grain, the sampling temporal grain can be considered.

Nevertheless, an other problem arises when the trend is computed. Usually, one compute the metric out of the data and considers the value to be representative of half a year or a year. In this case, there is no consensus on what should be the temporal grain: should one keep the grain of the sampling? Should one consider it to be the time-span covered by the sampling? In my analysis, the latter has been used (Table \@ref(tab:maintable)). 

For the cases where the metric was determined out of model [e.g. @harrison_assessing_2014], the final temporal grain was most of the time explicitly given, with an order of length of the year.

In short, temporal grain has no clear definition as it is different according to the metric computed, how it is computed and if the temporal trend is assessed. Even though it is now clear that the temporal grain of the sampling is important, there is no such consensus regarding the temporal grain of the trend.

Highest temporal grains are found with increasing spatial extent (Figure \ref{fig:spacetimegrain}). Data used in the selected article are mainly structured data, *i.e.* data following a well established sampling plan. This type of survey need resources and organizations which makes them sparser. Increasing the total area to sample makes *de facto* the time needed to sample the area to increase and eventually the temporal grain. However, this entanglement between spatial and temporal grain scales is only related to the constraints of the sampling. This limitation can be overcome thanks to citizen science data, which are nowadays more and more used [*e.g.* @bowler_winners_2021; @isaac_data_2020; @isaac_statistics_2014]. The opportunistic feature of these data make the number of observations more important, even over a large area. These data, with high temporal grain resolution for large spatial scales, are yet to be used to explore in more detail the temporal scaling of biodiversity trend.

```{r spacetimegrain, fig.cap = "Not sure about the relevance of this figure", fig.pos = "!H"}
# lol <- 
tab %>%
  mutate(`Spatial extent (Km²)` = as.numeric(unlist(tab$`Spatial extent (Km²)`)),
         `Temporal grain (year)`= as.numeric(unlist(tab$`Temporal grain (year)`))) %>% 
  select(`Spatial extent (Km²)`, `Temporal grain (year)`) %>%
  distinct() %>%
  ggplot(aes(`Spatial extent (Km²)`, `Temporal grain (year)`))+
  geom_point()+
  # geom_jitter()+
  expand_limits(x = 0)+
  geom_smooth(method = "lm", se = F)+
  scale_x_continuous(trans = "log10")+
  theme_classic()

# lm(`Temporal grain (year)`~`Spatial extent (Km²)`, data = lol)


# tab %>%
#   mutate(`Spatial extent (Km²)` = as.numeric(unlist(tab$`Spatial extent (Km²)`)),
#          `Temporal grain (year)`= as.numeric(unlist(tab$`Temporal grain (year)`))) %>% 
#   ggplot(aes(`Spatial grain (Km²)`, `Temporal grain (year)`))+
#   geom_jitter()+
#   geom_smooth(method = "lm", se = F)+
#   scale_x_discrete()+
#   theme_classic()
```


# Future directions

In order to find general patterns and reliable results about spatio-temporal scaling of biodiversity trends, replications are needed. That is, iteratively computing trends of biodiveristy at certain scales across the globe. However, out of `r length(unique(tab$Reference))` papers, 5 were located in North America and 12 in Europe. This is a striking example of the well-known lack of studies outside Europe and, to a lesser degree, North America [@fraixedas_state_2020]. Yet, biodiversity dynamic across scales in Europe is not representative of global dynamic, especially due to the lesser ecosystem representations, the variations in anthropization, conservation policies etc. Thus, studies replicating the computation of trends at several spatio-temporal scales are needed outside of Europe if one wants to understand the scaling of biodiverity dynamic. These non-european studies are even more needed when working at spatial grains the size or greater than Europe. In this case, spatial replication means having data about other continents.  

Even if a consesus exists about the importance of spatial grains, the final spatial grain of the metric computed is not always specified. One should consider the way the metric is computed, *e.g.* if it was summed, modeled, averaged over the sampling units. According to the method, the spatial grain can vary from the sampling unit to the sampling extent and even more if extrapolation methods are used. This may seem trivial but given the importance of spatial scaling of biodiveristy patterns [@storch_untangling_2004], one has to expect that it will be as important for its dynamic [@chase_species_2019]. By specifying systematically the spatial grain according to the metric used, we hope to facilitate the study of scaling biodiversity trends.

The importance of temporal scaling of biodiversity is known since @grinnell_role_1922, who used California birds to demonstrate the species-time relationship, which has since been proven to be common with other bird populations [@white_two-phase_2004]. Thus, as spatial grain, temporal grain is known to be important to explicit. However, there is no consensus on the definition of temporal grain and is thus specified in various ways: sometimes very precised [*e.g.* time of each census point, as in @schipper_contrasting_2016] and sometimes without explicit informations [*e.g.* @harrison_assessing_2014: *'The sites are visited twice a year (April to early May and late May to June), during which volunteers walk two parallel 1-km-long transect lines [...]'*]. The confusion becomes even more important when the trend of the metric is computed, as the temporal grain of the metric is considered to be representative of the unit time-span of the trend. In other words, when one computes and represent a trend on a figure, he usually use one point either every day, month or year. Thus, the temporal grain of the trend is different from the one of the computed metric, especially if one point on the figure embedds several census sessions. If one wants to study the temporal scaling of biodiversity trends, a clear definition of temporal grain for this context needs to be found. From the scientific literature, two definitions:

**1)** Temporal grain of the trend is the temporal grain of the computed metric. If it is so, it should be clearly specified

**2)** Temporal grain of the trend is the time span between census sessions/ time of computed metric. 

# Conclusion

Reviewing the scientific literature on this topic gives a glimpse of what still needs to be done to better understand the scaling of global avian biodiversity dynamic. As birds represent one of the most data furnished taxa of vertebrates, challenges about birds are just a part of what still need to be done for other taxa. 

<!-- for the bibliography -->
\newpage
\begin{singlespacing}
\printbibliography[heading=bibintoc, title={References}]
\end{singlespacing}


\newpage

# Supplementary materials {.unnumbered}

<!-- \renewcommand{\clearpage}{} -->
<!-- \vspace*{-1cm} -->

```{r notetable, echo=FALSE}
tab %>%
  select(Reference, `Spatial grain (Km²)`, Trend, Note) %>%
  as_tibble() %>%
  group_by(Reference) %>%
  collapse_rows_df(Reference) %>%
  ## escape = F is very important for the citation
  kbl(booktabs = T, escape = F, format = "latex", longtable = T,
      caption = "Supplementary informations about each article") %>%
  kableExtra::kable_styling(
   latex_options = c("hold_position", "repeat_header", "striped"),
   font_size = 10,
   full_width = FALSE
   ) %>%
  column_spec(1:3, width = "6.5em") %>%
  column_spec(4, width = "40em") %>% 
  landscape()
```

```{r, eval=F}
tab %>% 
    select(Metric, `Spatial grain (Km²)`, Trend) %>% 
    group_by(`Spatial grain (Km²)`) %>% 
    count(Trend) %>% 
    pivot_wider(1:3, names_from = Trend, values_from = n) %>% 
    mutate(n = sum(c(Decrease, Increase, Stable), na.rm = T)) %>% 
    mutate_at(c("Decrease", "Increase", "Stable"), funs(./n)) %>% ungroup() %>% 
    mutate_at(c("Decrease", "Increase", "Stable"), funs(format(round(., 2), nsmall = 2))) %>% 
    mutate_at(c("Decrease", "Increase", "Stable"), as.numeric) %>% 
  kable()
 
```

```{r, eval=F}
tmp <- 
  tab %>% 
    select(Metric, `Spatial grain (Km²)`, Trend) %>% 
    group_by(Metric, `Spatial grain (Km²)`) %>% 
    count(Trend) %>% 
    pivot_wider(1:4, names_from = Trend, values_from = n) %>% 
    mutate(n = sum(c(Decrease, Increase, Stable), na.rm = T)) %>% 
    mutate_at(c("Decrease", "Increase", "Stable"), funs(./n)) %>% ungroup() %>% 
    mutate_at(c("Decrease", "Increase", "Stable"), funs(format(round(., 2), nsmall = 2))) %>% 
    mutate_at(c("Decrease", "Increase", "Stable"), as.numeric)
    
  
  knitr::kable(tmp %>% 
                 filter(`Spatial grain (Km²)` == "Local") %>% 
                 # ungroup() %>%
                 select(-2),
               caption = "Local") %>% 
    column_spec(1:4, width = "6.5em") 
  
  knitr::kable(tmp %>% 
                 filter(`Spatial grain (Km²)` == "Regional") %>% 
                 # ungroup() %>%
                 select(-2),
               caption = "Regional") %>% 
    column_spec(1:4, width = "6.5em")
  
  knitr::kable(tmp %>% 
                 filter(`Spatial grain (Km²)` == "National") %>% 
                 # ungroup() %>%
                 select(-2),
               caption = "National") %>% 
    column_spec(1:4, width = "6.5em")
  
  knitr::kable(tmp %>% 
                 filter(`Spatial grain (Km²)` == "Global") %>% 
                 # ungroup() %>%
                 select(-2),
               caption = "Global") %>% 
    column_spec(1:4, width = "6.5em")  
```

